---
title: 'Stat 331 - Final'
author: "Lillie Allen, Nick Bukovec, Nikhil Koganti, Madeline Willett"
date: "3/2/2022"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  fig.align = "center"
  )

```

```{r libs}
library(tidyverse)
library(prettydoc)
library(here)
library(broom)
```

```{r formatting, include = FALSE}
# Center all plot titles
theme_update(plot.title = element_text(hjust = 0.5))
```

# Load Data
```{r load}
oil <- read_csv(here::here("oil_consumption_per_cap.csv"))
sdi <- read_csv(here::here("sdi.csv"))
```

# Clean Data

```{r clean}
oil_clean <- oil %>%
  # Coerce types
  transmute(
    country = as.factor(country),
    across(`1965`:`2019`, as.numeric)
    ) %>% 
  # Pivot 
  pivot_longer(
    cols = `1965`:`2019`, 
    names_to = "year",
    values_to = "Oil Consumption per Person (tonnes/person)",
    )

sdi_clean <- sdi %>% 
  # Coerce types
  transmute(
    country = as.factor(country),
    across(`1990`:`2019`, as.numeric)
    ) %>% 
  # Pivot
  pivot_longer(
    cols = `1990`:`2019`, 
    names_to = "year",
    values_to = "Sustainable Development Index",
    ) 

dataset <- sdi_clean %>% 
  # Join data by year and country
  inner_join(oil_clean, by = c("year", "country")) %>% 
  # Drop any missing data
  drop_na()
```

# Data Visualization

```{r}
dataset %>%
  ggplot(mapping = aes(
    x = `Oil Consumption per Person (tonnes/person)`, 
    y = `Sustainable Development Index`)) +
    geom_point() +
    geom_smooth(method = "lm") +
  labs(
    title = "Relationship Between Oil Consumption per Person and SDI",
    x = "Oil Consumption per Person (tonnes/person)",
    y = "Sustainable Development Index"
  )
```
 
# Visualization Over Time

```{r}
dataset %>%
  group_by(year) %>%
  mutate(ratio = (`Sustainable Development Index`/`Oil Consumption per Person (tonnes/person)`)) %>%
  summarize(mean = mean(ratio)) %>%
  ggplot(mapping = aes(
    x = `year`, 
    y = mean)) +
  geom_point() + 
  labs(
    title = "Mean Ratio of SDI to Oil Consumption Per Person Over Time",
    x = "Year",
    y = "Mean Ratio"
  ) +
  theme(axis.text.x = element_text(size = 7))
```

# Linear Regression

```{r}
dataset_lm <- dataset %>%
  lm(`Sustainable Development Index` ~ `Oil Consumption per Person (tonnes/person)`, data = .)
```

```{r}
broom::glance(dataset_lm) 
```
SDI = 67.32 -7.714(Oil Consumption)
The intercept, b0 = 67.32, is the average Sustainable Development Index y = SDI for those countries where the average Oil Consumption per Person is 0. It has no practical interpretation here, since observing an SDI of 0 is impossible. The slope for Oil Consumption per Person, b1 = -7.714,
summarizes the relationship between the SDI and Oil Consumption per Person variables. The sign is negative, suggesting a negative relationship between these two variables, meaning countries with higher average Oil Consumption per Person tend to have lower Sustainable Development Indexes. For every increase of 1 unit in Oil Consumption per Person, there is an associated decrease of, on average, 7.714 units of SDI.

# Residuals
```{r}
dataset_lm %>% 
  augment() %>% 
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```
