---
title: 'Stat 331 - Final'
author: "Lillie Allen, Nick Bukovec, Nikhil Koganti, Madeline Willett"
date: "3/2/2022"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  message = TRUE, 
  warning = TRUE, 
  fig.align = "center"
  )

```

```{r libs}
library(tidyverse)
library(prettydoc)
library(here)
library(broom)
library(ggtext)
library(gganimate)
library(kableExtra)
```

# Load Data
```{r load}
oil <- read_csv(here::here("oil_consumption_per_cap.csv"))
sdi <- read_csv(here::here("sdi.csv"))
```

# Clean Data

```{r clean}
oil_clean <- oil %>%
  # Coerce types
  transmute(
    country = as.factor(country),
    across(`1965`:`2019`, as.numeric)
    ) %>% 
  # Pivot 
  pivot_longer(
    cols = `1965`:`2019`, 
    names_to = "year",
    values_to = "Oil Consumption per Person (tonnes/person)",
    )

sdi_clean <- sdi %>% 
  # Coerce types
  transmute(
    country = as.factor(country),
    across(`1990`:`2019`, as.numeric)
    ) %>% 
  # Pivot
  pivot_longer(
    cols = `1990`:`2019`, 
    names_to = "year",
    values_to = "Sustainable Development Index",
    ) 

dataset <- sdi_clean %>% 
  # Join data by year and country
  inner_join(oil_clean, by = c("year", "country")) %>% 
  # Drop any missing data
  mutate(year = as.integer(year)) %>% 
  drop_na()
```

# Data Visualization

```{r}
# See Source 1
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#0072B2",
               "#D55E00")

dataset %>%
  mutate(
    country = as.factor(case_when(
      country == "Singapore" ~ "Singapore",
      country == "Bangladesh" ~ "Bangladesh",
      country == "Sri Lanka" ~ "Sri Lanka",
      country == "Uzbekistan" ~ "Uzbekistan",
      country == "United Arab Emirates" ~ "United Arab Emirates",
      country == "Kuwait" ~ "Kuwait",
      TRUE ~ "All Other Countries"
    ))
  ) %>% 
  ggplot(mapping = aes(
    x = `Oil Consumption per Person (tonnes/person)`, 
    y = `Sustainable Development Index`,
    color = country,
    size = case_when(
      country %in% c("Singapore", "Bangladesh", "Sri Lanka", "Uzbekistan", 
                     "United Arab Emirates", "Kuwait") ~ 3,
      TRUE ~ 0.5
    ) ,
    alpha = case_when(
      country %in% c("Singapore", "Bangladesh", "Sri Lanka", "Uzbekistan", 
                     "United Arab Emirates", "Kuwait") ~ 0.95,
      TRUE ~ 0.94
    ))) + 
  geom_jitter() +
  # See Source 2
  labs(
    title = "**How Oil Consumption Does Not Allow for Sustainable Development** 
    <br />
    <span style='font-size:11pt'>Countries such as<span style='color:#E69F00;'>
    Bangladesh</span>, <span style='color:#CC79A7;'>Sri Lanka</span>, and 
    <span style='color:#D55E00;'>Uzbekistan</span> shine while other 
    countries like <br /><span style='color:#009E73;'>Singapore</span>, the 
    <span style='color:#0072B2;'>United Arab Emirates</span>, and 
    <span style='color:#56B4E9;'>Kuwait</span> guzzle oil without 
    benefiting their citizens. </span>",
    x = "Oil Consumption per Person (tonnes/person)",
    y = "Sustainable Development Index",
    caption = "Data from Gapminder",
  ) +
  scale_y_continuous(limits = c(0, 110)) +
  scale_x_sqrt() +
  scale_color_manual(values = cbPalette, name = "") +
  scale_size_continuous(range = c(0.5, 1.3), name = "") +
  scale_alpha_continuous(range = c(0.45, 0.7), name = "Country") +
  # See Source 3
  theme(
    text = element_text(family = "Helvetica"),
    plot.title = element_markdown(lineheight = 1.1),
    plot.caption = element_text(hjust = 1, 
                                color = "blue", 
                                face = "italic"),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none")
```
 
# Visualization Over Time

```{r}
# See Source 1
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#0072B2",
               "#D55E00")
anim_plot <- dataset %>%
  ggplot(mapping = aes(
    x = `Oil Consumption per Person (tonnes/person)`, 
    y = `Sustainable Development Index`,
    size = `Oil Consumption per Person (tonnes/person)`,
    color = country
  )) +
  geom_jitter(
    alpha = 0.80,
    # size = 0.65
  ) +
  # See Source 2
  labs(
    title = "**Visualizing Oil Consumption and Sustainable Development ({frame_time})** 
    <br />
    <span style='font-size:11pt'>As countries consume more oil, the benefits
    to the lives of their citizens are outweighed <br /> by their carbon 
    footprints. However, as time has passed more countries have lowered their <br/>oil consumption while
    maintaining their sustainable development index.
    </span>",
    x = "Oil Consumption per Person (tonnes/person)",
    y = "Sustainable Development Index",
    caption = "Data from Gapminder",
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_sqrt() +
  # scale_color_manual(values = cbPalette, name = "") +
  # scale_colour_brewer() +
  scale_size_continuous(range = c(0.4, 10), name = "") +
  # See Source 3
  theme(
    text = element_text(family = "Helvetica"),
    plot.title = element_markdown(lineheight = 1.1),
    plot.caption = element_text(hjust = 1, 
                                color = "blue", 
                                face = "italic"),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none"
    ) +
  # https://gganimate.com/
  transition_time(year) +
  ease_aes(default = 'cubic-in-out') +
  # https://gganimate.com/articles/gganimate.html#labeling-1
  enter_fade() +
  exit_fade()
  
animate(anim_plot,
        duration = 20,
        fps = 10)

```

# Linear Regression

```{r}
dataset_lm <- dataset %>%
  lm(`Sustainable Development Index` ~ `Oil Consumption per Person (tonnes/person)`, data = .)
```

```{r}
# TODO: Format this equation in LaTex
broom::glance(dataset_lm) 
```

SDI = 67.32 -7.714 (Oil Consumption)

The intercept, b0 = 67.32, is the average Sustainable Development Index y = SDI for those countries where the average Oil Consumption per Person is 0. It has no practical interpretation here, since observing an SDI of 0 is impossible. The slope for Oil Consumption per Person, b1 = -7.714, summarizes the relationship between the SDI and Oil Consumption per Person variables. The sign is negative, suggesting a negative relationship between these two variables, meaning countries with higher average Oil Consumption per Person tend to have lower Sustainable Development Indexes. For every increase of 1 unit in Oil Consumption per Person, there is an associated decrease of, on average, 7.714 units of SDI.

```{r}
lm_tibble <- broom::augment(dataset_lm) %>% 
  summarise('Response Variance' = var(`Sustainable Development Index`),
            'Fitted Value Variance' = var(.fitted),
            'Residual Variance' = var(.resid)) %>% 
  mutate('Proportion of Variance Accounted for in Model' = `Fitted Value Variance` / (`Response Variance` + `Fitted Value Variance` + `Residual Variance`))
```

```{r}
lm_tibble %>%
  kbl() %>%
  kable_styling()
```

# Discussion
The variance in the response variable, 374.7564, tells us how much Countries differ in Sustainable Development Index. The variance in the fitted values, 150.5912, from our regression model tells us how much the fitted values from our linear regression model vary. The variance of the residuals, 224.1652, tells us how much “the left-overs” from the model vary. The proportion of the variability in the response values that was accounted for by our regression model is 0.2009188; thus, our linear regression model explains about 20% of the variation in Sustainable Development Index. This suggests that our model does not tell us everything we need to know about Sustainable Development Index.
